Bayesian Statistical Modeling with Predictors from LLMs

### 2024-05-27 04-paper/paper-LLM-probmod.tex
corrected passage on length-normalization

\paragraph{LLM scores \& predictions.}
The most obvious \emph{item-level score} an (autoregressive) LLM provides for each choice option $y_{kl}$ is its log-probability:\footnote{
  More elaborate item-level scores include corrections for variable length of answer options \citep[e.g.,][]{BrownMann2020:Language-Models} or variation in base rate among answer options \citep[e.g.,][]{HoltzmanWest2021:Surface-Form-Co}.
  From the point of view of experimental psychology, these corrections are \emph{post hoc} fixes to improperly balanced experimental materials.
  For the purposes of this paper, where answer options are all equally long and commensurable, these corrections may be temporarily ignored for simplicity.
  \mf{@PT: check whether this is correct also after new LLaMA results.}
}
%
\begin{align*}
  % S\left( y_{ki}, x_{k} \right) =
  \text{S}_{kl}
  % = \log P_{\text{LLM}} \left(y_{ki} \mid x_{k} \right)
  =  \sum_{i=1}^{|y_{kl}|} \log P_{\text{LLM}} \left(w_{kli} \mid x_{k}, w_{kl1}, \dots, w_{kl(i-1)} \right)  \,.
\end{align*}
%

>>>>

\paragraph{LLM scores \& predictions.}
The most obvious \emph{item-level score} an (autoregressive) LLM provides for each choice option $y_{kl}$ is its log-probability:\footnote{
  Commonly used item-level scores include corrections for variable length of answer options \citep[e.g.,][]{BrownMann2020:Language-Models} or variation in base rate among answer options \citep[e.g.,][]{HoltzmanWest2021:Surface-Form-Co}.
  Following common practice in the current literature, we also use length-corrected, average log-probabilities as raw scores for all options.
  As most options have identical number of tokens, this only affects the numerical values of softmax-parameter $\alpha$, as introduced below.}
}
%
\begin{align*}
  % S\left( y_{ki}, x_{k} \right) =
  \text{S}_{kl}
  % = \log P_{\text{LLM}} \left(y_{ki} \mid x_{k} \right)
  =  \sum_{i=1}^{|y_{kl}|} \log P_{\text{LLM}} \left(w_{kli} \mid x_{k}, w_{kl1}, \dots, w_{kl(i-1)} \right)  \,.
\end{align*}
%
